---
title: "Hypothesis Testing HTML"
author: "Anton"
date: "3/3/2020"
output:
  word_document: default
  html_document: default
---

Set-up Libraries
```{r include=FALSE}
library(magrittr)
library(lubridate)
library(reshape)
library(tidyverse)
library(plyr)
library(readr)
library(readxl)
library(car)
library(psych)
library(lsr)
library(BSDA)
library(PASWR)
library(Rmisc)

```

Read/Upload Data
```{r include=FALSE}
F_D = read.csv("C:/Users/Anton/OneDrive - Singapore Management University/Python and R/R/R_Group_Project/Working_Data/Full_Climate_Data.csv")
S_D = read.csv("C:/Users/Anton/OneDrive - Singapore Management University/Python and R/R/R_Group_Project/Working_Data/Singapore_Aggregate_Data.csv")
```

Examine/View Data
```{r include=FALSE}
# View(S_D)
# View(F_D)
```
# Hypothesis_testing

# Tempreture

## Wrangle Tempreture data into chunks necessary for analysis
Making subset for Mean_Tempreture date only
```{r}
F_T = S_D[,c(1,6)]
# View(F_T)
```

Omit all N/A from Tempreture data
```{r}
F_T_O = na.omit(F_T)
row.names(F_T_O) = NULL
# View(F_T_O)
```

Make two subset for test
```{r}
T1 = F_T_O[1:6955,]
# View(T1)
```
```{r}
T2 = F_T_O[6956:13910,]
# View(T2)
```

## Check for normality our Tempreture data
```{r}
shapiro.test(T1$Average.of.Mean_Temperature[1:5000])
```
The data is not normally distributed

Vizulize distribition
```{r}
hist(T1$Average.of.Mean_Temperature, breaks = 250, col = "purple", main = "Singapore Mean Tempreture 1982-2001", xlab = "Mean tempreture")
```
```{r}
shapiro.test(T2$Average.of.Mean_Temperature[1:5000])
```
For the second chunk the data distributed not normal either
Vizulize distribition
```{r}
hist(T2$Average.of.Mean_Temperature, breaks = 270, col = "purple", main = "Singapore Mean Tempreture 2001-2020", xlab = "Mean tempreture")
```

## Run the non-parametrick paired test for Tempreture
Now we need to use non-parametrick test to check either or not the data have same symmetrical distribution
```{r}
wilcox.test(T1$Average.of.Mean_Temperature,T2$Average.of.Mean_Temperature,paired = TRUE,alternative = "two.sided")
```
The distribution of t1 and t2 are significally different, that can account for climate change

# Rainfall
## Wrangle Data
Making subset for Mean_Rainfall date only
```{r}
F_R = S_D[,c(1,2)]
row.names(F_R) = NULL
# View(F_R)
```

Check for N/A
```{r}
count(is.na(F_R))
```

Make two subset for test
```{r}
R1 = F_R[1:7320,]
# View(R1)
```
```{r}
R2 = F_R[7321:14640,]
# View(R2)
```

## Check for normality our Rain data
```{r}
shapiro.test(R1$Average.of.Daily.Rainfall.Total..mm.[1:5000])
```
The data is not normally distributed
```{r}
hist(R1$Average.of.Daily.Rainfall.Total..mm., breaks = 250, col = "purple", main = "Singapore Mean Rainfall 1980-2000", xlab = "Mean Rainfall")
```
```{r}
hist(log(R1$Average.of.Daily.Rainfall.Total..mm.), breaks = 250, col = "purple", main = "Singapore Log Mean Rainfall 1980-2000", xlab = "Log Mean Rainfall")
```



```{r}
shapiro.test(R2$Average.of.Daily.Rainfall.Total..mm.[1:5000])
```
The data is not normally distributed
```{r}
hist(R2$Average.of.Daily.Rainfall.Total..mm., breaks = 250, col = "purple", main = "Singapore Mean Rainfall 2000-2020", xlab = "Mean Rainfall")
```
```{r}
hist(log(R2$Average.of.Daily.Rainfall.Total..mm.), breaks = 250, col = "purple", main = "Singapore Log Mean Rainfall 2000-2020", xlab = "Log Mean Rainfall")
```

## Run the non-parametrick paired test for Rain
Now we need to use non-parametrick test to check either or not the data have same symmetrical distribution
```{r}
wilcox.test(R1$Average.of.Daily.Rainfall.Total..mm.,R2$Average.of.Daily.Rainfall.Total..mm.,paired = TRUE,alternative = "two.sided")
```
The p-value is right on 1% level significance, we reject Ho. Rain distribution for 20 years are different, which account for climate change


# Wind

## Wrangle Wind data into chunks necessary for analysis
Making subset for Mean_Tempreture date only
```{r}
F_W = S_D[,c(1,9)]
# View(F_W)
```

Omit all N/A from Tempreture data
```{r}
F_W_O = na.omit(F_W)
row.names(F_W_O) = NULL
# View(F_W_O)
```

Make two subset for test
```{r}
W1 = F_W_O[1:7136,]
# View(W1)
```
```{r}
W2 = F_W_O[7137:14272,]
# View(W2)
```

## Check for normality our Wind data
```{r}
shapiro.test(W1$Average.of.Mean.Wind.Speed..km.h.[1:5000])
```
The data distribution is not normal
```{r}
hist(W1$Average.of.Mean.Wind.Speed..km.h., breaks = 300, col = "purple", main = "Singapore Mean Wind 1981-2000", xlab = "Mean Wind")
```

```{r}
shapiro.test(W2$Average.of.Mean.Wind.Speed..km.h.[1:5000])
```
The Data distributed not normall

```{r}
hist(W2$Average.of.Mean.Wind.Speed..km.h., breaks = 250, col = "purple", main = "Singapore Mean Wind 2000-2020", xlab = "Mean Wind")
```

## Run the non-parametrick paired test for Rain
Now we need to use non-parametrick test to check either or not the data have same symmetrical distribution
```{r}
wilcox.test(W1$Average.of.Mean.Wind.Speed..km.h.,W2$Average.of.Mean.Wind.Speed..km.h.,paired = TRUE,alternative = "two.sided")
```
we reject Ho hence there is a change in wind distributions as well 



































